ISYE 6501 | Submission HW2

### 2022-09-07 hw2/submission.Rmd
final ques 3.1

```{r, include=FALSE}

@@

2.  Choose i'th part as test data and train knn on the rest 4

    1.  Find the threshold which maximises AUC using ROC curve

@@

4.  When all parts are rotated into test set, calculate the measure for accuracy of model:

@@

    2.  Min test accuracy across the 5 folds

@@

The above summary shows the mean of R1 and relative size of each fold. Ideally for all the folds this should be 45% mean and 20%, but we see minor deviations due to random sampling.

```{r}
#Function to train model and return prediction on test
train_test_i = function(df_train,df_test, k0 = 5){
  f = R1~A1+A2+A3+A8+A9+A10+A11+A12+A14+A15
  model = kknn(f,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
  return (fitted(model))
}
```

@@

### Step 4: Iterate on different models and use CV for best model

```{r, include=FALSE}
#Iterate over k and get mean test accuracy

end_k = 77 #half of test set size

@@

We have also kept standard deviation, such that we can able to infer the variance of accuracy across cross folds. This is helpful because the maximum CV accuracy for different values across k, varies within a percent (See graph below).

#### Plot the accuracy with k, and find optimal k which maximizes accuracy

```{r}
ggplot(data = acc_df, aes(x = k_kknn, y = mean_cv_accuracy)) +
  geom_line() +geom_point() +
  geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
             color ="magenta", size = 1.5, alpha = 0.5) +
  ggtitle(paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))

@@

However, as noted above the nuance to note her is that, for k \> 30, that accuracy mostly oscillates at almost the same range with delta of 0.5. But since the standard deviation of max accuracy among cross folds is 2.45 itself, this choice of k becomes a little tricky. In other words, we

@@

### Part 2: Train, Validation and Test on knnn

```{r}
#split data into 70% train, 15% validation and 15% test
df_copy2 = org_cc_data
df_copy2$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)

@@

```{r, include=FALSE}
acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,end_k,2))))
acc_df2$k_kknn = seq(3,end_k,2)
acc = lapply(seq(3,end_k,2), function(x) 
  train_valid_func(df_train = df_train, df_test = df_valid, k_choice = x))
acc_df2$acc_validation_set = unlist(acc)
```

@@

Here we see that for k = 23, we get maximum average test accuracy of 84%.

Note, that both the optimal k value here and the optimal accuracy value are different than part 1 above. The test accuracy here is lower at 84% compared to 87% of accuracy in CV case, which is expected, since 84% is accuracy at true test, while cv accuracy is averaged across all the folds.

>>>>

```{r}

@@

2.  Choose i'th part as test data and train knn model on the rest 4 parts

    1.  Find the optimal threshold using the ROC curve

@@

4.  When all parts are rotated into the test set, calculate the measure for accuracy of model:

@@

    2.  Min test accuracy across the 5 folds (optional)

@@

The above summary shows the mean of R1 and relative size of each fold. Ideally for all the folds this should be 45% mean and 20% proportion, but we see minor deviations in individual folds due to random sampling.

```{r}
#Function to train model and return prediction probabilities on test
train_test_i = function(df_train,df_test, k0 = 5){
  f = R1~A1+A2+A3+A8+A9+A10+A11+A12+A14+A15
  model = kknn(f,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
  return (fitted(model))
}

# decide best threshold
best_threshold = function(pred_prob, df_test=org_cc_data){
  myroc <- roc(df_test$R1,as.numeric(unlist(pred_prob)), smooth=FALSE)
  print(paste("AUC is",auc(myroc)))
  #plot(myroc, auc.polygon=TRUE)
  return (coords(myroc,x="best")["threshold"][[1]])
}

#Accuracy for best threshold
acc_func = function(y_act, pred_probs, thresh = 0.5){
  y_pred = lapply(pred_probs, function(x) if(x>=thresh) 1 else 0)
  y_pred = as.numeric(unlist(y_pred))
  acc_overall = sum(y_act == y_pred)/length(y_act)
  acc_overall = round(acc_overall,4)* 100
  return (acc_overall)
}
```

@@

### Step 4: Iterate on different models(i.e on different values of k) and use CV for best model

Here we differentiate two knn models, only using the k hyperparameter

```{r, results='hide'}
#Iterate over k and get mean test accuracy

end_k = 77 #almost half of test set size

@@

We have also kept standard deviation, such that we will be able to infer the variance of accuracy across cross folds. This is helpful because the maximum CV accuracy for different values across k varies in a very close interval (around 87.5 %) -\> see graph below

#### Plot the accuracy with k (hyperparameter of knn), and find optimal k which maximizes accuracy

```{r}
ggplot(data = acc_df, aes(x = k_kknn, y = mean_cv_accuracy)) +
  geom_line() +geom_point() +
  geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
             color ="blue", size = 1.5, alpha = 0.5) +
  ggtitle(paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))

@@

However, as noted above the nuance to note here is that, for k \> 30, accuracy mostly oscillates at almost the same range with delta of 0.5. But since the standard deviation of max accuracy among cross folds is 2.45 itself, this choice of k becomes a little tricky. In other words, any k value above 30 seems appropriate, as the performance separation between the models is minimal

@@

### Part 2: Train, Validation and Test on k-knn

```{r}
set.seed(77)

#split data into 70% train, 15% validation and 15% test
df_copy2 = org_cc_data
df_copy2$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)

@@

```{r, results='hide'}
acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,end_k,2))))
acc_df2$k_kknn = seq(3,end_k,2)
acc = lapply(seq(3,end_k,2), function(x) 
  train_valid_func(df_train = df_train, df_test = df_valid, k_choice = x))
acc_df2$acc_validation_set = unlist(acc)
```

@@

Note, that both the optimal k value here and the optimal accuracy value are different than part 1 above.

The test accuracy here is the true (or best available) evaluation of the model chosen by us.

However one might also argue that the optimal k values from both the approaches lie in a similar range, since any k value \>30 can be chosen from cross validation too.
